\documentclass[../main.tex]{subfiles}
 
\begin{document}

\begin{thebibliography}{99}

\bibitem{Manninen} S. M. Reimann and M. Manninen. Electronic structure of quantum dots. \emph{Rev. Mod.
Phys.}, \textbf{74}:1238, 2002.

\bibitem{QDotSolar} Zerui Zheng, Haining Ji, Peng Yu and Zhiming Wang: Recent Progress Towards Quantum Dot Solar Cells with Enhanced Optical Absorption. \emph{Nanoscale Res Lett} 2016, \textbf{11}:266.

\bibitem{QDotSolar2} Xinzheng Lan, Oleksandr Voznyy, F. Pelayo García de Arquer, Mengxia Liu, Jixian Xu, Andrew H. Proppe, Grant Walters, Fengjia Fan, Hairen Tan, Min Liu, Zhenyu Yang, Sjoerd Hoogland, and Edward H. Sargent. \textbf{10.6\% Certified Colloidal Quantum Dot Solar Cells via Solvent-Polarity-Engineered Halide Passivation}. \emph{Nano Letters} \textbf{2016} 16 (7), 4630-4634. DOI: 10.1021/acs.nanolett.6b01957.

\bibitem{QDotMed} Shan Jin, Yanxi Hu, Zhanjun Gu, Lei Liu, and Hai-Chen Wu, "Application of Quantum Dots in Biological Imaging," \emph{Journal of Nanomaterials}, vol. 2011, Article ID 834139, 13 pages, 2011. doi:10.1155/2011/834139.

\bibitem{QDotMed2} Bentolila LA, Michalet X, Pinaud FF, Tsay JM, Doose S, Li JJ, Sundaresan G, Wu AM, Gambhir SS, Weiss S. Quantum dots for molecular imaging and cancer medicine. Discov. Med. 2005;5:213–218.

\bibitem{QDotLaser} Ledentsov, N.N. Quantum dot laser. \emph{Semiconductor Science and Technology} \textbf{2011}, 26, 014001.

\bibitem{QDotLaser2} Veena Hariharan Iyer, Rekha Mahadevu, and Anshu Pandey. \textbf{Low Threshold Quantum Dot Lasers}. \emph{The Journal of Physical Chemistry Letters} \textbf{2016} 7 (7), 1244-1248. DOI: 10.1021/acs.jpclett.6b00430.

\bibitem{QDotComp} Hai-Rui Wei and Fu-Guo Deng. Scalable quantum computing based on stationary spin qubits in coupled quantum dots inside double-sided optical microcavities. \emph{Scientific Reports} \textbf{4}, , Article number: 7551 (2014). doi:10.1038/srep07551.

\bibitem{QDotComp2} Eriksson, M.A., Friesen, M., Coppersmith, S.N. et al. Quantum Information Processing (2004) 3: 133. doi:10.1007/s11128-004-2224-z

\bibitem{QMC1} F. Bolton. Fixed-phase quantum Monte Carlo method applied to interacting electrons in
a quantum dot. Phys. Rev. B \textbf{54}:4780 (1996). DOI: \url{https://doi.org/10.1103/PhysRevB.54.4780}

\bibitem{QMC2} Francesco Pederiva, C. J. Umrigar, and E. Lipparini. Diffusion Monte Carlo study of
circular quantum dots. Phys. Rev. B \textbf{62}:8120 (2000). DOI: \url{https://doi.org/10.1103/PhysRevB.62.8120}

\bibitem{QMC3} A. D. Güçlü, Jian-Sheng Wang, and Hong Guo. Disordered quantum dots: A diffusion
quantum Monte Carlo study. Phys. Rev. B \textbf{68}:035304 (2003). DOI: \url{https://doi.org/10.1103/PhysRevB.68.035304}

\bibitem{QDotBenchmarks} M.~.L.~Pedersen, G.~Hagen, M.~Hjorth-Jensen, S.~Kvaal,  and F.~Pederiva. Ab initio computation of the energies of circular quantum dots. Phys.~Rev.~B {\bf 84}, 115302 (2011). DOI: \url{https://doi.org/10.1103/PhysRevB.84.115302}

\bibitem{FYS4411-CG} M.~Hjorth-Jensen, {\em Conjugate gradient methods and other optimization methods},  \url{https://github.com/CompPhysics/ComputationalPhysics2/blob/master/doc/pub/cg/pdf/cg-print.pdf}, Read: 17.03.2016.

\bibitem{Taut} M.~Taut, \emph{Two electrons in an external oscillator potential: Particular analytic solutions
of a Coulomb correlation problem}, Phys. Rev. A {\bf 48}, 3561 - 3566 (1993).

\bibitem{FYS4411-Slides} M.~Hjorth-Jensen, \emph{Computational Physics 2: Variational Monte Carlo methods}, \url{https://github.com/CompPhysics/ComputationalPhysics2/blob/master/doc/pub/vmc/pdf/vmc-print.pdf}, Read: 16.06.2016.

\bibitem{FYS4411-LectureNotes} M.~Hjorth-Jensen, \emph{Computational Physics Lecture Notes Fall 2015},  \url{https://github.com/CompPhysics/ComputationalPhysics2/blob/master/doc/Literature/lectures2015.pdf}, Chapter 16.10, Read: 16.06.2016.

\bibitem{Griffiths} David j. Griffiths, \emph{Introduction to Quantum Mechanics}, Pearson,
2nd edition edition, 2005.

\bibitem{Scientific Computing} Aslak Tveito, Hans Petter Langtangen, Bjørn Frederik Nielsen and Xing Cai, \emph{Elements of Scientific Computing}, Springer, 1st edition, 2010.

\bibitem{Hermite} B.H. Bransden and C.J. Joachain, \emph{Physics of Atoms and molecules}, Longman, 1983. Chapter
2.

\bibitem{Schrodinger} Georigia State University, \emph{Time Independent Schrodinger Equation}, Hyperphysics entry, \url{http://hyperphysics.phy-astr.gsu.edu/hbase/quantum/Scheq.html}, Read 23.05.17

\bibitem{CP1_Project 2} M.~Hjorth-Jensen, \url{https://github.com/CompPhysics/ComputationalPhysics/blob/master/doc/Projects/2015/Project2/project2_2015.pdf}

\bibitem{master project} M.~Hjorth-Jensen, \url{https://github.com/CompPhysics/ThesisProjects/blob/master/doc/pub/quantumdots/pdf/quantumdots-minted.pdf}

\bibitem{basicMB} M.~Hjorth-Jensen, \emph{Hartree-Fock methods and introduction to Many-body Theory}, \url{https://github.com/CompPhysics/ComputationalPhysics2/blob/gh-pages/doc/pub/basicMB/pdf/basicMB-print.pdf}

\bibitem{Ruder} Sebastian Ruder, \emph{An overview of gradient descent optimization
algorithms}, 2016, 	\href{https://arxiv.org/abs/1609.04747}{arXiv:1609.04747 [cs.LG]}

%\bibitem{QD-wiki} \url{https://en.wikipedia.org/wiki/Quantum_dot}, Read: 16.06.2016.

\bibitem{Stroustrup} Bjarne Stroustrup, \emph{Bjarne Stroustrup's C++ Glossary},  \url{http://www.stroustrup.com/glossary.html#Gpolymorphism}, Read: 09.05.2017.

\bibitem{gccDocs} \emph{Using the GNU Compiler Collection (GCC)}, section 3.10: \emph{Options That Control Optimization}, \url{http://gcc.gnu.org/onlinedocs/gcc/Optimize-Options.html}, Read 10.05.2017

\bibitem{github} Christian Fleishcer, Github Repository 2017,  \url{https://github.com/christianfleischer/Quantum-Dot-Project}.

\bibitem{Blaise} Blaise Barney, Lawrence Livermore National Laboratory, \emph{Introduction to Parallel Computing}, \url{https://computing.llnl.gov/tutorials/parallel_comp/}, Read 26.05.2017

\bibitem{Wheeler} David A. Wheeler, \emph{Secure Programming HOWTO}, Chapter 7,  \url{https://www.dwheeler.com/secure-programs/Secure-Programs-HOWTO/avoid-race.html}%\url{https://en.wikipedia.org/wiki/Race_condition}, Read 23.05.2017

\bibitem{Lin} Charles Lin, \emph{Logic Gates}, computer organization class notes, 2003, \url{https://www.cs.umd.edu/class/sum2003/cmsc311/Notes/Comb/gates.html} %\url{https://en.wikipedia.org/wiki/Logic_gate}, Read 23.05.2017

\bibitem{Yang} Yang Min Wang. Coupled-Cluster Studies of Double Quantum
Dots. Master’s thesis, University of Oslo, 2011.

\bibitem{Jorgen} Jørgen Høgberget. Quantum Monte-Carlo Studies of Generalized Many-body Systems. 
Master’s thesis, University of Oslo, 2013.

\bibitem{Reimann} Sarah Reimann. Quantum-mechanical systems in traps and Similarity Renormalization Group theory. Master’s thesis, University of Oslo, 2013.

\bibitem{Hirth} Christoffer Hirth. Studies of quantum dots: Ab initio coupledcluster analysis using OpenCL and GPU programming. Master’s thesis, University of Oslo, 2012.

\bibitem{Olsen} Veronica K. B. Olsen. Full Configuration Interaction Simulation of Quantum Dots. Master’s thesis, University of Oslo, 2012.

\bibitem{Armadillo} Conrad Sanderson and Ryan Curtin. Armadillo: a template-based C++ library for linear algebra. 
Journal of Open Source Software, Vol. 1, pp. 26, 2016. \url{http://dx.doi.org/10.21105/joss.00026}

\bibitem{Open MPI} Edgar Gabriel, Graham E. Fagg, George Bosilca, Thara Angskun, Jack J. Dongarra, Jeffrey M. Squyres, Vishal Sahay, Prabhanjan Kambadur, Brian Barrett, Andrew Lumsdaine, Ralph H. Castain, David J. Daniel, Richard L. Graham, and Timothy S. Woodall. Open MPI: Goals, Concept, and Design of a Next Generation MPI Implementation. In Proceedings, 11th European PVM/MPI Users' Group Meeting, Budapest, Hungary, September 2004.

\bibitem{Matplotlib} John D. Hunter. Matplotlib: A 2D Graphics Environment, Computing in Science \& Engineering, 9, 90-95 (2007), \href{http://dx.doi.org/10.1109/MCSE.2007.55}{DOI:10.1109/MCSE.2007.55}, Publisher link:  \url{http://aip.scitation.org/doi/abs/10.1109/MCSE.2007.55}

\bibitem{NumPy} Stéfan van der Walt, S. Chris Colbert and Gaël Varoquaux. The NumPy Array: A Structure for Efficient Numerical Computation, Computing in Science \& Engineering, 13, 22-30 (2011), \href{http://dx.doi.org/10.1109/MCSE.2011.37}{DOI:10.1109/MCSE.2011.37}, Publisher link:  \url{http://aip.scitation.org/doi/abs/10.1109/MCSE.2011.37}

\bibitem{SymPy} Meurer A, Smith CP, Paprocki M, Čertík O, Kirpichev SB, Rocklin M, Kumar A, Ivanov S, Moore JK, Singh S, Rathnayake T, Vig S, Granger BE, Muller RP, Bonazzi F, Gupta H, Vats S, Johansson F, Pedregosa F, Curry MJ, Terrel AR, Roučka Š, Saboo A, Fernando I, Kulal S, Cimrman R, Scopatz A. (2017) SymPy: symbolic computing in Python. PeerJ Computer Science 3:e103 \url{https://doi.org/10.7717/peerj-cs.103}

\bibitem{Multiwells} Daniel V. Schroeder, \emph{Multiple Wells} \url{http://physics.weber.edu/schroeder/quantum/MultipleWells.pdf}, Read 26.05.2017



\end{thebibliography}
\end{document}